---
title: "Reproducible Research: Peer Assessment 1"
author: "Mark Drobnis"
date: "December 9, 2016"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown file(rmd). R Markdown is a simple formatting syntax that 
can be used to produce HTML, PDF, and MS Word documents. 
For more details use the following link: 
<http://rmarkdown.rstudio.com>.

by clicking the **Knit** button a html document will be generated that includes standard output and embedded R code chunks within the document.

This document makes use of data from personal activity monitoring devices. 
These devices collects data every 5 minute through out the day. 
The data contains two months of data for every 5 minute interval durning the 
months of October and November of 2012 and number of steps taken for each interval.

The data for this file should be downloaded along with rmd file:
<https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip>
Place all file in a new directory: 

##Reproducible Research: Peer Assessment 1

get needed libraries

```{r}
library(ggplot2)
library(tcltk)
library(Hmisc)
library(sqldf)
```

##Code for reading in the dataset and/or processing the data

###1. Load the data (i.e. read.csv())

```{r}
##Make this match the directory you placed the files in
setwd('C:/Users/I63706/Documents/data/RRWeekPeerAssignment/repdataFdataFactivity/')
if(file.exists('activity.csv')){
    aData <- read.csv('activity.csv')}
aData$dateinterval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", aData$interval), format='%H:%M')
BDsteps <- tapply(aData$steps, aData$date, sum, na.rm=TRUE)
```

###2. Make a histogram of the steps taken each day

```{r}
    qplot(BDsteps, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

###3. Calculate the mean steps taken per 5 minute interval

```{r}
    
    Meanstepspi <-  tapply(aData$steps, as.numeric(aData$dateinterval), mean, na.rm=TRUE)
    MeanSteps <- mean(aData$steps, na.rm=TRUE)
    Medianstepspi <- tapply(aData$steps, as.numeric(aData$dateinterval), median, na.rm=TRUE)
    Mediansteps <- median(aData$steps, na.rm=TRUE )
```

###4. Make a time series plot of the mean per 5 minute intervals.

```{r echo}
    maverageStepsBlock <- aggregate(x=list(meanSteps=aData$steps), by=list(interval=aData$interval), FUN=mean, na.rm=TRUE)
    ggplot(data=maverageStepsBlock, aes(x=interval, y=meanSteps)) + geom_line() + xlab("5-minute interval") +
     ylab("average number of steps taken") 
    
```

###4a. Make a time series plot of the median per 5 minute intervals.


```{r}
    averageStepsBlock <- aggregate(x=list(medianSteps=aData$steps), by=list(interval=aData$interval), FUN=median, na.rm=TRUE)
    ggplot(data=averageStepsBlock, aes(x=interval, y=medianSteps)) + geom_line() + xlab("median number of steps taken") +
    ylab("5-minute interval") 
```

###5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
MaxstepsInterval <- sqldf("select interval from maverageStepsBlock where meanSteps = (select max(meanSteps) from maverageStepsBlock)")
MaxstepsInterval
```

###6. Code to describe and show a strategy for imputing missing data

```{r}
aDataI <- aData
aDataI$steps <- impute(aData$steps, fun=mean)
```

###7. Make a histogram of the total number of steps taken each day

```{r}
sByDayI <- tapply(aDataI$steps, aDataI$date, sum)
qplot(sByDayI, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
ByDayMeanI <- mean(sByDayI)
ByDayMedianI <- median(sByDayI)
```

###8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
aDataI$dateType <-  ifelse(as.POSIXlt(aDataI$date)$wday %in% c(0,6), 'weekend', 'weekday')
aDataI <- aggregate(steps ~ interval + dateType, data=aDataI, mean)
ggplot(aDataI, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("average number of steps taken")
```
